Concurrent Programming in Harmony

### 2022-01-06 book/paper.tex
Added a selection of operations on strings

\section*{Exercises}
\begin{problems}
\item Add the line <{print(x, triangle(x))}> to the end of the
program and create an output png file.  Before you look at it, what
do you think it should look like?
\item See what happens if, instead of initializing <{result}> to 0,
you initialize it to 1.  (You do not need to understand the error report at this time.  They will be explained in more detail in \autoref{ch:harmonymachine}.)
\item Write a Harmony program that computes squares by repeated adding.  So, the program
should compute the square of <{x}> by adding <{x}> to an initial value of 0 <{x}> times.
\end{problems}

@@

@@

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
<{e}> & <{e}> is an expression\\
\hline
<{lv = [lv =]... e}> & <{lv}> is an lvalue and <{e}> is an expression\\
\hline
<{lv} [op]= e}> & <{op}> is one of \texttt{+}, \texttt{-},
\texttt{*}, \texttt{/}, \texttt{//}, \texttt{\%},
\texttt{\&}, \texttt{|}, \texttt{\^}, <{and}>, <{or}>\\
\hline
<{l: S}> & <{l}> is a label \\
\hline
<{assert b [, e]}> & <{b}> is a boolean.  Optionally report value of expression <{e}>\\
\hline
<{await b}> & <{b}> is a boolean\\
\hline
<{const a = e}> & <{a}> is a bound variable, <{e}> is a constant expression\\
\hline
<{def m a: S}> & <{m}> is an identifier, <{a}> a bound variable\\
\hline
<{del lv}> & delete <{lv}>\\
\hline
<{for a[:b] in e [where c]: S}> & <{a}> and <{b}> are bound variables, <{e}> is a set or dictionary \\
\hline
<{from m import ...}> & <{m}> identifies a module \\
\hline
<{go c e}> & <{c}> is a context, <{e}> is an expression \\
\hline
<{if b: S else: S}> & <{b}> is a boolean, <{S}> is a list of statements\\
\hline
<{import m, ...}> & <{m}> identifies a module \\
\hline
<{invariant e}> & <{e}> is an invariant\\
\hline
<{let a = e: S}> & <{a}> is a bound variable, <{e}> is an expression\\
\hline
<{pass}> & do nothing\\
\hline
<{print e>}> & <{e}> is an expression \\
\hline
<{sequential v, ...}> & <{v}> has sequential consistency \\
\hline
<{spawn [eternal] m e [, t]}> & <{m}> is a method,
<{e}> is an expression, <{t}> is the thread-local state \\
\hline
<{trap m e}> & <{m}> is a method and <{e}> is an expression \\
\hline
<{var v = e}> & <{v}> is a new variable, <{e}> is an expression\\
\hline
<{when b: S}> & <{b}> is a boolean, <{S}> a list of statements\\
\hline
<{when exists a in e: S}> & <{a}> is a bound variable, <{e}> is an expression\\
\hline
<{while b: S}> & <{b}> is a boolean, <{S}> a list of statements\\
\hline
\end{tabular}

@@

The statement <{for x in y: S1; S2; ...}>
iterates over <{y}> and executes for each element the statements
<{S1, S2, ...}>.
<{y}> must be a set or a dictionary.  <{y}> is evaluated only once at
the beginning of the evaluation of this statement.  In case of a set,
the result is sorted (using Harmony's global order on all values).
In case of a dictionary, the statement iterates over the values of
the dictionary in the order of the keys.  This makes iterating over
lists intuitive and identical to Python.
For each element, the statements <{S1, S2, ...}> are executed with
local variable <{y}> having the value of the element.
<{x}> can be a pattern such as $(a)$ or $(a, (b, c))$.
If the pattern cannot be matched, Harmony detects and error.
It is allowed, but discouraged, to assign different values to <{x}>
within statements <{S1, S2, ...}>.

>>>>

\section*{Exercises}
\begin{problems}
\item Write a Harmony program that uses <{choose}> instead of <{spawn}> to create the same
output DFA as \autoref{fig:threads}(b).
\item Add the line <{print(x, triangle(x))}> to the end of the
program and create an output png file.  Before you look at it, what
do you think it should look like?
\item See what happens if, instead of initializing <{result}> to 0,
you initialize it to 1.  (You do not need to understand the error report at this time.  They will be explained in more detail in \autoref{ch:harmonymachine}.)
\item Write a Harmony program that computes squares by repeated adding.  So, the program
should compute the square of <{x}> by adding <{x}> to an initial value of 0 <{x}> times.
\end{problems}

@@

Native operations on strings include the following:

\begin{center}
\begin{tabular}{|l|l|}
\hline
<{s k}> & indexing \\
\hline
<{s + s + ...}> & concatenation \\
\hline
<{s * n}> & <{n}> copies of <{s}> concatenated \\
\hline
<{v [not] in s}> & check if <{v}> is [not] a substring in <{s}> \\
\hline
<{len s}> & the length of <{s}> \\
\hline
<{str e}> & string representation of any value <{e}> \\
\hline
\end{tabular}
\end{center}

@@

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
<{e}> & <{e}> is an expression\\
\hline
<{lv = [lv =]... e}> & <{lv}> is an lvalue and <{e}> is an expression\\
\hline
<{lv} [op]= e}> & <{op}> is one of \texttt{+}, \texttt{-},
\texttt{*}, \texttt{/}, \texttt{//}, \texttt{\%},
\texttt{\&}, \texttt{|}, \texttt{\^}, <{and}>, <{or}>\\
\hline
<{l: S}> & <{l}> is a label \\
\hline
<{assert b [, e]}> & <{b}> is a boolean.  Optionally report value of expression <{e}>\\
\hline
<{await b}> & <{b}> is a boolean\\
\hline
<{const a = e}> & <{a}> is a bound variable, <{e}> is a constant expression\\
\hline
<{def m a: S}> & <{m}> is an identifier, <{a}> a bound variable\\
\hline
<{del lv}> & delete <{lv}>\\
\hline
<{for a[:b] in e [where c]: S}> & <{a}> and <{b}> are bound variables, <{e}> is a set, dictionary, or string \\
\hline
<{from m import ...}> & <{m}> identifies a module \\
\hline
<{go c e}> & <{c}> is a context, <{e}> is an expression \\
\hline
<{if b: S else: S}> & <{b}> is a boolean, <{S}> is a list of statements\\
\hline
<{import m, ...}> & <{m}> identifies a module \\
\hline
<{invariant e}> & <{e}> is an invariant\\
\hline
<{let a = e: S}> & <{a}> is a bound variable, <{e}> is an expression\\
\hline
<{pass}> & do nothing\\
\hline
<{print e>}> & <{e}> is an expression \\
\hline
<{sequential v, ...}> & <{v}> has sequential consistency \\
\hline
<{spawn [eternal] m e [, t]}> & <{m}> is a method,
<{e}> is an expression, <{t}> is the thread-local state \\
\hline
<{trap m e}> & <{m}> is a method and <{e}> is an expression \\
\hline
<{var v = e}> & <{v}> is a new variable, <{e}> is an expression\\
\hline
<{when b: S}> & <{b}> is a boolean, <{S}> a list of statements\\
\hline
<{when exists a in e: S}> & <{a}> is a bound variable, <{e}> is an expression\\
\hline
<{while b: S}> & <{b}> is a boolean, <{S}> a list of statements\\
\hline
\end{tabular}

@@

The statement <{for x in y: S1; S2; ...}>
iterates over <{y}> and executes for each element the statements
<{S1, S2, ...}>.
<{y}> must be a set, dictionary, or string.  <{y}> is evaluated only once at
the beginning of the evaluation of this statement.  In case of a set,
the result is sorted (using Harmony's global order on all values).
In case of a dictionary, the statement iterates over the values of
the dictionary in the order of the keys.  This makes iterating over
lists intuitive and identical to Python.
For each element, the statements <{S1, S2, ...}> are executed with
local variable <{y}> having the value of the element.
<{x}> can be a pattern such as $(a)$ or $(a, (b, c))$.
If the pattern cannot be matched, Harmony detects and error.
It is allowed, but discouraged, to assign different values to <{x}>
within statements <{S1, S2, ...}>.
